### 2018-07-05 submission/manuscript.Rmd
fix abstract

# Read in GC data
gc_data <- read_csv("../data/process/tables/v4_gc_wilcox_test_summary.csv")

@@

**Results.** Although the total counts for specific taxa varied by polymerase, we found no noticeable differences in total counts based on high and low GC content (P-value $\geqslant$ `r round(min(pull(gc_data, p.value)), digits = 2)`). Chimera prevalence in mock communities varied by polymerase with differences being most notable at 35 cycles (Kappa = `r round((filter(error_summary, cycles == "35x", taq == "K") %>% pull(chimera_prevalence_median))*100, digits = 2)`% (median) versus Platinum = `r round((filter(error_summary, cycles == "35x", taq == "PL") %>% pull(chimera_prevalence_median))*100, digits = 2)`%) and this variation persisted after chimera removal using VSEARCH. We also observed positive correlations between chimera prevalence and the number of OTUs with Platinum having the highest (R^2^ = `r max(pull(r2_table, r2))`) and Kappa having the worst (R^2^ = `r min(pull(r2_table, r2))`). When analyzing mock community samples the variation in the number of OTUs detected by the polymerases was observable as early as 20 cycles (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`). There also was a large range in the number of OTUs amplified by the polymerases at 35 cycles (Accuprime = `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion = `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). When analyzing fecal samples we observed that the range in the number of OTUs detected was not consistent between HiFi DNA polymerases (eg. at 35 cycles Accuprime = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` (min - max) versus Phusion = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). Additionally, the median number of OTUs varied by HiFi DNA polymerase used at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001). Random Forest models were most successful at classifying individual donor samples rather than polymerase or number of cycles used (P-value $\leqslant$ `r format(max(filter(dunn_rf_data, str_detect(var1, "fecal"), str_detect(var2, "fecal")) %>% pull(bh)), scientific = T, digits = 3)`). Additionally, the most important OTUs in the polymerase and number of cycle models were not the most important in the individual donor sample model.

**Conclusions.** Although there are 16S rRNA gene sequencing differences based on polymerase and number of cycles used, they are small with respect to the biological differences between indivdiuals. Additionally, our observations were not able to identify GC-based differences in amplification based on the number of cycles used. Collectively, these results provide evidence that a real biological difference between groups, based on 16S rRNA gene sequencing, should be detectable regardless of polymerase and number of cycles used.

@@

***Statistical analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data transformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. Differences in the total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the comparison of the number of OTUS in fecal samples the data was normalized to each individual by cycle number to account for the biological variation. Bray-Curtis distance matrices were generated using mothur. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and Kruskal-Wallis tests within R. The Random Forest models were run using the caret package (v `r paste(packageVersion("caret")[1])`) [@caret_citation]. A total of 100 10-fold CV runs on different 80/20 splits of the data was run to generate a range of the Logloss value. The probability of a correct call was obtained from this Logloss value by taking the negative natural logrithim. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable, correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

@@

>>>>

# Read in GC data
gc_data <- read_csv("../data/process/tables/v4_gc_wilcox_test_summary.csv") %>% 
  filter(cycles != "fifteen") %>% 
  mutate(bh = p.adjust(p.value, method = "BH"))

gc_genome_data <- read_csv("../data/process/tables/whole_genome_gc_wilcox_test_summary.csv") %>% 
  filter(cycles != "fifteen") %>% 
  mutate(bh = p.adjust(p.value, method = "BH"))

gc_summary <- read_csv("../data/process/tables/gc_summary_table.csv")

@@

**Results.** In addition to the total counts for specific taxa varying by polymerase, we found noticeable differences in relative abundance based on high and low GC content (P-value $\leqslant$ `r round(max(filter(gc_data, bh < 0.05) %>% pull(bh)), digits = 2)`). Chimera prevalence in mock communities varied by polymerase with differences being most notable at 35 cycles (Kappa = `r round((filter(error_summary, cycles == "35x", taq == "K") %>% pull(chimera_prevalence_median))*100, digits = 2)`% (median) versus Platinum = `r round((filter(error_summary, cycles == "35x", taq == "PL") %>% pull(chimera_prevalence_median))*100, digits = 2)`%) and this variation persisted after chimera removal using VSEARCH. We also observed positive correlations between chimera prevalence and the number of OTUs with Platinum having the highest (R^2^ = `r max(pull(r2_table, r2))`) and Kappa having the worst (R^2^ = `r min(pull(r2_table, r2))`). When analyzing mock community samples the variation in the number of OTUs detected by the polymerases was observable as early as 20 cycles (P-value = `r format(filter(mock_anova_summary, sub_sample_level == 1000, cycle == "20x")[, "bh"], digits = 1, nsmall = 1)`). There also was a large range in the number of OTUs amplified by the polymerases at 35 cycles (Accuprime = `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` versus Phusion = `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(mock_num_otu_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). When analyzing fecal samples we observed that the range in the number of OTUs detected was not consistent between HiFi DNA polymerases (eg. at 35 cycles Accuprime = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Accuprime") %>% pull(taq_max)` (min - max) versus Phusion = `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_min)` - `r filter(fecal_numOTU_data, cycles == "35x", taq == "Phusion") %>% pull(taq_max)`). Additionally, the median number of OTUs varied by HiFi DNA polymerase used at 35 cycles (P-value `r ifelse(max(filter(fs_anova_summary, cycle == "35x")[, "bh"]) < 0.0001, invisible("<"), invisible("check data"))` 0.0001). Random Forest models were most successful at classifying individual donor samples rather than polymerase or number of cycles used (P-value $\leqslant$ `r format(max(filter(dunn_rf_data, str_detect(var1, "fecal"), str_detect(var2, "fecal")) %>% pull(bh)), scientific = T, digits = 3)`). Additionally, the most important OTUs in the polymerase and number of cycle models were not the most important in the individual donor sample model.

**Conclusions.** Although there are 16S rRNA gene sequencing differences based on polymerase and number of cycles used, they are small with respect to the biological differences between indivdiuals. Collectively, these results provide evidence that a real biological difference between groups, based on 16S rRNA gene sequencing, should be detectable regardless of polymerase and number of cycles used.

@@

***Statistical analysis:***
All analysis was done with the R (v `r paste(version$major, ".", version$minor, sep = "")`) software package [@r_citation_2017]. Data transformation and graphing was completed using the tidyverse package (v `r paste(packageVersion("tidyverse")[1])`) and colors selected using the viridis package (v `r paste(packageVersion("viridis")[1])`) [@tidyverse_2017; @viridis_citation_2017]. High and low GC content was determined based on the median GC percentage of either the V4 region or the whole genome of the bacterial species used in the mock community. Differences in the total number of OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the comparison of the number of OTUS in fecal samples the data was normalized to each individual by cycle number to account for the biological variation. Bray-Curtis distance matrices were generated using mothur. The distance matrix data was analyzed using PERMANOVA with the vegan package (v `r paste(packageVersion("vegan")[1])`) [@vegan_citation] and Kruskal-Wallis tests within R. The Random Forest models were run using the caret package (v `r paste(packageVersion("caret")[1])`) [@caret_citation]. A total of 100 10-fold CV runs on different 80/20 splits of the data was run to generate a range of the Logloss value. The probability of a correct call was obtained from this Logloss value by taking the negative natural logrithim. For both error and chimera analysis, samples were tested using Kruskal-Wallis with a Dunns post-hoc test. Where applicable, correction for multiple comparison utilized the Benjamini-Hochberg method [@benjamini_controlling_1995].

@@

***16S rRNA gene sequencing counts do not differ by GC content:*** Although the 16S rRNA gene sequencing counts varied by polymerase used, there was no difference between V4 16S rRNA genes with high and low GC content at every amplification cycle [Figure 1 & Table S1]. Additionally, there was also no difference in 16S rRNA gene sequencing counts between high and low GC content genomes [Table S2].

### 2018-07-05 submission/manuscript.tex
fix abstract

\textbf{Results.} Although the total counts for specific taxa varied by
polymerase, we found no noticeable differences in total counts based on
high and low GC content (P-value \(\geqslant\) 0.23). Chimera prevalence
in mock communities varied by polymerase with differences being most
notable at 35 cycles (Kappa = 5.71\% (median) versus Platinum = 26.62\%)
and this variation persisted after chimera removal using VSEARCH. We
also observed positive correlations between chimera prevalence and the
number of OTUs with Platinum having the highest (R\textsuperscript{2} =
0.974) and Kappa having the worst (R\textsuperscript{2} = 0.478). When
analyzing mock community samples the variation in the number of OTUs
detected by the polymerases was observable as early as 20 cycles
(P-value = 0.002). There also was a large range in the number of OTUs
amplified by the polymerases at 35 cycles (Accuprime = 15 - 20 versus
Phusion = 14 - 73). When analyzing fecal samples we observed that the
range in the number of OTUs detected was not consistent between HiFi DNA
polymerases (eg. at 35 cycles Accuprime = 84 - 106 (min - max) versus
Phusion = 84 - 136). Additionally, the median number of OTUs varied by
HiFi DNA polymerase used at 35 cycles (P-value \textless{} 0.0001).
Random Forest models were most successful at classifying individual
donor samples rather than polymerase or number of cycles used (P-value
\(\leqslant\) 5.49e-07). Additionally, the most important OTUs in the
polymerase and number of cycle models were not the most important in the
individual donor sample model.

\textbf{Conclusions.} Although there are 16S rRNA gene sequencing
differences based on polymerase and number of cycles used, they are
small with respect to the biological differences between indivdiuals.
Additionally, our observations were not able to identify GC-based
differences in amplification based on the number of cycles used.
Collectively, these results provide evidence that a real biological
difference between groups, based on 16S rRNA gene sequencing, should be
detectable regardless of polymerase and number of cycles used.

@@

\textbf{\emph{Statistical analysis:}} All analysis was done with the R
(v 3.4.4) software package (R Core Team, 2017). Data transformation and
graphing was completed using the tidyverse package (v 1.2.1) and colors
selected using the viridis package (v 0.4.1) (Garnier, 2017; Wickham,
2017). Differences in the total number of OTUs were analyzed using an
ANOVA with a tukey post-hoc test. For the comparison of the number of
OTUS in fecal samples the data was normalized to each individual by
cycle number to account for the biological variation. Bray-Curtis
distance matrices were generated using mothur. The distance matrix data
was analyzed using PERMANOVA with the vegan package (v 2.4.5) (Oksanen
et al., 2017) and Kruskal-Wallis tests within R. The Random Forest
models were run using the caret package (v 6.0.78) (Jed Wing et al.,
2017). A total of 100 10-fold CV runs on different 80/20 splits of the
data was run to generate a range of the Logloss value. The probability
of a correct call was obtained from this Logloss value by taking the
negative natural logrithim. For both error and chimera analysis, samples
were tested using Kruskal-Wallis with a Dunns post-hoc test. Where
applicable, correction for multiple comparison utilized the
Benjamini-Hochberg method (Benjamini \& Hochberg, 1995).

@@

>>>>

\textbf{Results.} In addition to the total counts for specific taxa
varying by polymerase, we found noticeable differences in relative
abundance based on high and low GC content (P-value \(\leqslant\) 0.04).
Chimera prevalence in mock communities varied by polymerase with
differences being most notable at 35 cycles (Kappa = 5.71\% (median)
versus Platinum = 26.62\%) and this variation persisted after chimera
removal using VSEARCH. We also observed positive correlations between
chimera prevalence and the number of OTUs with Platinum having the
highest (R\textsuperscript{2} = 0.974) and Kappa having the worst
(R\textsuperscript{2} = 0.478). When analyzing mock community samples
the variation in the number of OTUs detected by the polymerases was
observable as early as 20 cycles (P-value = 0.002). There also was a
large range in the number of OTUs amplified by the polymerases at 35
cycles (Accuprime = 15 - 20 versus Phusion = 14 - 73). When analyzing
fecal samples we observed that the range in the number of OTUs detected
was not consistent between HiFi DNA polymerases (eg. at 35 cycles
Accuprime = 84 - 106 (min - max) versus Phusion = 84 - 136).
Additionally, the median number of OTUs varied by HiFi DNA polymerase
used at 35 cycles (P-value \textless{} 0.0001). Random Forest models
were most successful at classifying individual donor samples rather than
polymerase or number of cycles used (P-value \(\leqslant\) 5.49e-07).
Additionally, the most important OTUs in the polymerase and number of
cycle models were not the most important in the individual donor sample
model.

\textbf{Conclusions.} Although there are 16S rRNA gene sequencing
differences based on polymerase and number of cycles used, they are
small with respect to the biological differences between indivdiuals.
Collectively, these results provide evidence that a real biological
difference between groups, based on 16S rRNA gene sequencing, should be
detectable regardless of polymerase and number of cycles used.

@@

\textbf{\emph{Statistical analysis:}} All analysis was done with the R
(v 3.4.4) software package (R Core Team, 2017). Data transformation and
graphing was completed using the tidyverse package (v 1.2.1) and colors
selected using the viridis package (v 0.4.1) (Garnier, 2017; Wickham,
2017). High and low GC content was determined based on the median GC
percentage of either the V4 region or the whole genome of the bacterial
species used in the mock community. Differences in the total number of
OTUs were analyzed using an ANOVA with a tukey post-hoc test. For the
comparison of the number of OTUS in fecal samples the data was
normalized to each individual by cycle number to account for the
biological variation. Bray-Curtis distance matrices were generated using
mothur. The distance matrix data was analyzed using PERMANOVA with the
vegan package (v 2.4.5) (Oksanen et al., 2017) and Kruskal-Wallis tests
within R. The Random Forest models were run using the caret package (v
6.0.78) (Jed Wing et al., 2017). A total of 100 10-fold CV runs on
different 80/20 splits of the data was run to generate a range of the
Logloss value. The probability of a correct call was obtained from this
Logloss value by taking the negative natural logrithim. For both error
and chimera analysis, samples were tested using Kruskal-Wallis with a
Dunns post-hoc test. Where applicable, correction for multiple
comparison utilized the Benjamini-Hochberg method (Benjamini \&
Hochberg, 1995).

@@

***16S rRNA gene sequencing counts do not differ by GC \url{content:***}
Although the 16S rRNA gene sequencing counts varied by polymerase used,
there was no difference between V4 16S rRNA genes with high and low GC
content at every amplification cycle {[}Figure 1 \& Table S1{]}.
Additionally, there was also no difference in 16S rRNA gene sequencing
counts between high and low GC content genomes {[}Table S2{]}.
